Time invariant analysis of epidemics with \pkg{EpiCompare}

### 2021-03-25 jss/EpiCompare-paper/EpiCompare-paper.Rmd
some updates to section 3's final parts.

Although this package does not focus on fitting a model to data, we do provide some flexible functions for simulation of basic discrete-time epidemic-state models.  These functions produce individual level information and can be naturally combined with `agents_to_aggregate()` to view these simulations through an aggregate lens.  The function `simulate_SIR_agents()` simulates an SIR epidemic with user inputs for the number of simulations, the initial number in each state, the infection and recovery parameters $(\beta, \gamma$), and the total number of discrete time steps.  This function allows for easy access to SIR model analysis and comparison.  Beyond SIR models, the function `simulate_agents()` takes as input a user-specified transition matrix and other epidemic parameters to allow the user to create simulations of an outbreak for \textit{any} number of states and any number of transitions among them.  This flexibility in states can be used to also reflect group-based dynamics.  This allows for users to explore the space of models in an intuitive way without getting bogged down by too much mathematical detail. For consistency, we have made output from `simulate_agents()` and `simulate_SIR_agents()` compatible with `agents_to_aggregate()` so aggregate information may easily be accessed.

@@

Post-processing of modeling and simulation consists of making summary statistics, plots, tables, and other ways to disseminate information to the public.  For example, comma separated value files (`.csv`) are a  standard way to share information within tables.  However, model output is often far more complicated than what a traditional `.csv` would allow.  \shannon{Do we need three sentences about csvs?} As a result, a number of epidemic modeling packages return a special class, specific to their modeling.  The special classes often contain a plethora of information from residuals, model diagnostics, input parameters, and more.  While incredibly useful, these special classes can be difficult for novice coders to navigate.

To this end, we have adapted a series of `fortify`-style functions, called `fortify_aggregate()` which transform output from packages like \pkg{pomp} and \pkg{EpiModel} into tidy-styled data frames which contain information about the total number of individuals in each state at a given time, for a given simulation.  These fortify functions have output that is consistent with that of `agents_to_aggregate()`.

\shannon{Here Ben talks about filament compression and tidy_dist mats, etc}

@@

>>>>

\ben{[Ben says: think about this section and if it highlights that we can bring in outside models...]}

\ben{After getting a good sense of what a past or current epidemice looks like through EDA, the next step is often model fitting and/or simulations. In this step, and the next step (post-processing), we discuss how to include practioners' use of tools to create models and simulations that exist outside of \pkg{EpiCompare}.} This package does not focus on fitting a model to data, we do provide some flexible functions for simulation of basic discrete-time epidemic-state models. \ben{After estimating or predicting transition rates between states,} these functions produce individual level \ben{simulations\sout{information}} and can be \ben{\sout{naturally}} combined with `agents_to_aggregate()` to view these simulations through an aggregate lens.  The function `simulate_SIR_agents()` simulates an SIR epidemic with user inputs for the number of simulations, the initial number in each state, the infection and recovery parameters $(\beta, \gamma$), and the total number of discrete time steps.  This function allows for easy access to SIR model analysis and comparison.  Beyond SIR models, the function `simulate_agents()` takes as input a user-specified transition matrix and other epidemic parameters to allow the user to create simulations of an outbreak for \textit{any} number of states and any number of transitions among them.  This flexibility in states can be used to also reflect group-based dynamics.  This allows for users to explore the space of models in an intuitive way without getting bogged down by too much mathematical detail. For consistency, we have made output from `simulate_agents()` and `simulate_SIR_agents()` compatible with `agents_to_aggregate()` so aggregate information may easily be accessed.

@@

\ben{[Ben says: I think we should remind the reader that we care more about simulations, in order to compare fitted models between themselves and the true epidemics.]}

\ben{[Ben says: this replaces the first paragraph below]}
\ben{If the practitioner wishes to compare between models and with their true epidemics, one needs to process their models and simulations.} \ben{In general, [p}ost-processing of modeling and simulation consists of making summary statistics, plots, tables, and other ways to disseminate information to the public.]  \ben{The summaries can be very complex, and a}[s a result, a number of epidemic modeling \ben{\proglang{R}} packages return a special class, specific to their modeling.  The special classes often contain a plethora of information from residuals, model diagnostics, input parameters, and more.  While incredibly useful, these special classes can be difficult for novice coders to navigate.]

\ben{[Ben says: replaced with above paragraph]} Post-processing of modeling and simulation consists of making summary statistics, plots, tables, and other ways to disseminate information to the public. For example, comma separated value files (`.csv`) are a  standard way to share information within tables.  However, model output is often far more complicated than what a traditional `.csv` would allow.  \shannon{Do we need three sentences about csvs?} As a result, a number of epidemic modeling packages return a special class, specific to their modeling.  The special classes often contain a plethora of information from residuals, model diagnostics, input parameters, and more.  While incredibly useful, these special classes can be difficult for novice coders to navigate.

To this end, we have \ben{provide\sout{adapted}} a series of `fortify`-style \ben{methods\sout{functions}}\footnote{\ben{[Ben says: Not tied to this change.]}}, called `fortify_aggregate()` which transform output from packages like \pkg{pomp} and \pkg{EpiModel} into tidy-styled data frames which contain information about the total number of individuals in each state at a given time, for a given simulation.  These fortify functions have output that is consistent with that of `agents_to_aggregate()`.

\shannon{[Shannon says: Here Ben talks about filament compression and tidy_dist mats, etc.]} \ben{[Ben says: not currently thinking about including tidy_dist stuff - will need to think about this as it is inlcudding in the pipeline 2 image...]}

\ben{[Ben says: new paragraph]} To utilize simulations epidemics in later time-invariant analysis we also provide a function to convert temporally defined epidemics to filamental representations. Specifically, we provide the function `filament_compression` to convert simulation(s) to filaments as expressed by presenting the epidemic as a ordering of some common fixed number of points so that they are equally spaced along the original path in the proportional state space\footnote{\ben{[Ben: this should be cleaned up.]}}.

@@

\ben{[This still needs to be worked on - just a reminder.]}

### 2021-03-25 jss/EpiCompare-paper/EpiCompare-paper.tex
some updates to section 3's final parts.

Although this package does not focus on fitting a model to data, we do
provide some flexible functions for simulation of basic discrete-time
epidemic-state models. These functions produce individual level
information and can be naturally combined with
\texttt{agents\_to\_aggregate()} to view these simulations through an
aggregate lens. The function \texttt{simulate\_SIR\_agents()} simulates
an SIR epidemic with user inputs for the number of simulations, the
initial number in each state, the infection and recovery parameters
\((\beta, \gamma\)), and the total number of discrete time steps. This
function allows for easy access to SIR model analysis and comparison.
Beyond SIR models, the function \texttt{simulate\_agents()} takes as
input a user-specified transition matrix and other epidemic parameters
to allow the user to create simulations of an outbreak for \textit{any}
number of states and any number of transitions among them. This
flexibility in states can be used to also reflect group-based dynamics.
This allows for users to explore the space of models in an intuitive way
without getting bogged down by too much mathematical detail. For
consistency, we have made output from \texttt{simulate\_agents()} and
\texttt{simulate\_SIR\_agents()} compatible with
\texttt{agents\_to\_aggregate()} so aggregate information may easily be
accessed.

@@

Post-processing of modeling and simulation consists of making summary
statistics, plots, tables, and other ways to disseminate information to
the public. For example, comma separated value files (\texttt{.csv}) are
a standard way to share information within tables. However, model output
is often far more complicated than what a traditional \texttt{.csv}
would allow. \textcolor{orange}{Do we need three sentences about csvs?}
As a result, a number of epidemic modeling packages return a special
class, specific to their modeling. The special classes often contain a
plethora of information from residuals, model diagnostics, input
parameters, and more. While incredibly useful, these special classes can
be difficult for novice coders to navigate.

To this end, we have adapted a series of \texttt{fortify}-style
functions, called \texttt{fortify\_aggregate()} which transform output
from packages like \pkg{pomp} and \pkg{EpiModel} into tidy-styled data
frames which contain information about the total number of individuals
in each state at a given time, for a given simulation. These fortify
functions have output that is consistent with that of
\texttt{agents\_to\_aggregate()}.

@@

>>>>

\textcolor{violet}{[Ben says: think about this section and if it highlights that we can bring in outside models...]}

\textcolor{violet}{After getting a good sense of what a past or current epidemice looks like through EDA, the next step is often model fitting and/or simulations. In this step, and the next step (post-processing), we discuss how to include practioners' use of tools to create models and simulations that exist outside of \pkg{EpiCompare}.}
This package does not focus on fitting a model to data, we do provide
some flexible functions for simulation of basic discrete-time
epidemic-state models.
\textcolor{violet}{After estimating or predicting transition rates between states,}
these functions produce individual level
\textcolor{violet}{simulations\sout{information}} and can be
\textcolor{violet}{\sout{naturally}} combined with
\texttt{agents\_to\_aggregate()} to view these simulations through an
aggregate lens. The function \texttt{simulate\_SIR\_agents()} simulates
an SIR epidemic with user inputs for the number of simulations, the
initial number in each state, the infection and recovery parameters
\((\beta, \gamma\)), and the total number of discrete time steps. This
function allows for easy access to SIR model analysis and comparison.
Beyond SIR models, the function \texttt{simulate\_agents()} takes as
input a user-specified transition matrix and other epidemic parameters
to allow the user to create simulations of an outbreak for \textit{any}
number of states and any number of transitions among them. This
flexibility in states can be used to also reflect group-based dynamics.
This allows for users to explore the space of models in an intuitive way
without getting bogged down by too much mathematical detail. For
consistency, we have made output from \texttt{simulate\_agents()} and
\texttt{simulate\_SIR\_agents()} compatible with
\texttt{agents\_to\_aggregate()} so aggregate information may easily be
accessed.

@@

\textcolor{violet}{[Ben says: I think we should remind the reader that we care more about simulations, in order to compare fitted models between themselves and the true epidemics.]}

\textcolor{violet}{[Ben says: this replaces the first paragraph below]}

\textcolor{violet}{If the practitioner wishes to compare between models and with their true epidemics, one needs to process their models and simulations.}
\textcolor{violet}{In general, [p}ost-processing of modeling and
simulation consists of making summary statistics, plots, tables, and
other ways to disseminate information to the public.{]}
\textcolor{violet}{The summaries can be very complex, and a}{[}s a
result, a number of epidemic modeling \textcolor{violet}{\proglang{R}}
packages return a special class, specific to their modeling. The special
classes often contain a plethora of information from residuals, model
diagnostics, input parameters, and more. While incredibly useful, these
special classes can be difficult for novice coders to navigate.{]}

\textcolor{violet}{[Ben says: replaced with above paragraph]}
Post-processing of modeling and simulation consists of making summary
statistics, plots, tables, and other ways to disseminate information to
the public. For example, comma separated value files (\texttt{.csv}) are
a standard way to share information within tables. However, model output
is often far more complicated than what a traditional \texttt{.csv}
would allow. \textcolor{orange}{Do we need three sentences about csvs?}
As a result, a number of epidemic modeling packages return a special
class, specific to their modeling. The special classes often contain a
plethora of information from residuals, model diagnostics, input
parameters, and more. While incredibly useful, these special classes can
be difficult for novice coders to navigate.

To this end, we have \textcolor{violet}{provide\sout{adapted}} a series
of \texttt{fortify}-style
\textcolor{violet}{methods\sout{functions}}\footnote{\textcolor{violet}{[Ben says: Not tied to this change.]}},
called \texttt{fortify\_aggregate()} which transform output from
packages like \pkg{pomp} and \pkg{EpiModel} into tidy-styled data frames
which contain information about the total number of individuals in each
state at a given time, for a given simulation. These fortify functions
have output that is consistent with that of
\texttt{agents\_to\_aggregate()}.

\textcolor{violet}{[Ben says: new paragraph]} To utilize simulations
epidemics in later time-invariant analysis we also provide a function to
convert temporally defined epidemics to filamental representations.
Specifically, we provide the function \texttt{filament\_compression} to
convert simulation(s) to filaments as expressed by presenting the
epidemic as a ordering of some common fixed number of points so that
they are equally spaced along the original path in the proportional
state
space\footnote{\textcolor{violet}{[Ben: this should be cleaned up.]}}.

@@

\textcolor{violet}{[This still needs to be worked on - just a reminder.]}
